
# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r}
setwd("~/Git/RepData_PeerAssessment1")
library(plyr)
fit_orig = read.csv("activity.csv")

fit1 = fit_orig[fit_orig$steps != 0 & !is.na(fit_orig$steps),]
fit1$date = as.Date(fit1$date)

fit2 = ddply(fit1,'date',function(x) c(obs_count = nrow(x),tot_steps = sum(x$steps),avg_steps = mean(x$steps), median_steps = median(x$steps)))
fit3 = ddply(fit1,'interval',function(x) c(avg_steps = mean(x$steps)))

```

## Histogram showing number of total steps per day (y-axis is number/frequency of days) 

```{r fig.width=7, fig.height=6}
library(lattice)

histogram(fit2$tot_steps,type="count",xlab = "Steps per Day",ylab = "Frequency (Days)")

options(digits = 2)
avg_steps = as.character(round(mean(fit2$tot_steps),2))
median_steps = as.character(median(fit2$tot_steps))
```

### The Mean number of steps per day is `r avg_steps`.
### The Median number of steps per day is `r median_steps`.

## What is the average daily activity pattern?  Average Number of Steps Taken per 5-minute Interval

```{r}

xyplot(avg_steps~interval,type="l",data=fit3,ylab = "Average Number of Steps", xlab = "Interval in Minutes")
```

## Which interval on average has the daily maximum number of steps?
```{r}
maxint = fit3[fit3$avg_steps == max(fit3$avg_steps),]
max_steps_interval = maxint$interval
max_avg_steps = maxint$avg_steps
```

### Interval `r max_steps_interval` has the maximum average number of steps of all intervals (`r max_avg_steps`)

## Imputing missing values

```{r}
na_val_count = nrow(fit_orig[is.na(fit_orig$steps),])
```
### There are `r na_val_count` missing (NA) values in the original data set.

# Creates a dataset that fills in the NA values **with the mean for 5-minute interval that corresponds to that of the missing interval.   <<< MY MISSING DATA STRATEGY :) **
```{r}
fit4 = fit_orig
fit4$date = as.Date(fit4$date)

# Function that assigns all rows with steps = NA to average number of steps for that interval (ignore NA's)
fn = function(x) { if (is.na(x$steps))  return(mean(fit_orig[fit_orig$interval == x$interval,]$steps,na.rm = TRUE)) else return(x$steps) }

fit4 = ddply(fit4,.(date,interval),.fun = fn)

names(fit4)[3] = "steps"

fit5 = ddply(fit4,'date',function(x) c(obs_count = nrow(x),tot_steps = sum(x$steps),avg_steps = mean(x$steps), median_steps = median(x$steps)))
```

## After computation for missing values, Histogram showing number of total steps per day (y-axis is number/frequency of days) 

```{r fig.width=7, fig.height=6}

histogram(fit5$tot_steps,type="count",xlab = "Steps per Day",ylab = "Frequency (Days)")

options(digits = 2)
avg_steps_narm = as.character(round(mean(fit5$tot_steps),0))
median_steps_narm = as.character(round(median(fit5$tot_steps),0))
```

### The Mean number of steps per day after missing values for steps calculated is `r avg_steps_narm`.
### The Median number of steps per day after missing values for steps calculated is `r median_steps_narm`.

### It appears that the missing data strategy that created the simulated data set (find average for the 5-minute interval) produces similar results to the mean and median calculations in the original data set when missing values are ignored.

## What is the average daily activity pattern?  Average Number of Steps Taken per 5-minute Interval

```{r}

## Are there differences in activity patterns between weekdays and weekends?

fit6 = ddply(fit4,.(date,interval,steps),.fun = function(x) {if (weekdays(x$date) == "Saturday" | weekdays(x$date) =="Sunday")  return("Weekend")  else return("Weekday")})
names(fit6)[4] = "dayofweek"
fit6$dayofweek = as.factor(fit6$dayofweek)


xyplot(steps~interval | dayofweek,type="l",data=fit6,ylab = "Average Number of Steps", xlab = "Interval in Minutes",layout = c(1,2))

